---
title       : Introduction to my Class Project
subtitle    : Plotting 2013 Census Data by County
author      : Lynda Finn
job         : Coursera Developing Data Products Course Project
framework   : io2012        # {io2012, html5slides, shower, dzslides, ...}
highlighter : highlight.js  # {highlight.js, prettify, highlight}
hitheme     : tomorrow      # 
widgets     : []            # {mathjax, quiz, bootstrap}
mode        : selfcontained # {standalone, draft}
knit        : slidify::knit2slides
logo        : my_logo.png
---

## Web Application to Map some of the 2013 Census Data by County

 * The web app is found at http://lyndafinn.shinyapps.io/Maps  

 * The census data is from 2013. I chose some of the most interesting variables  for plotting. It can be found here: http://quickfacts.census.gov/qfd/download_data.html

 * The application was based on the the shiny tutorial lesson 5, but I altered it significantly. The original tutorial can be seen here: http://shiny.rstudio.com/tutorial/lesson5/

--- .class #id 

## Modifications to Tutorial

In my code, I made the following modifications to the shiny tutorial:

* I used a wider set of census variables from a more up to date data source
* I modified the map coloring to be in the heatmap style
* I improved and corrected the labeling of the map categories. 
 * Each category now represents 20% of the data. This was necessary to plot such variables as population, because of its extreme skewness. Without the modification, the map was almost the lowest category (all yellow) 
  * In the original, the categories were improperly labeled with cutpoints, not ranges or midpoints
* I generalized the approach so that variables other than percentages could be plotted

--- .class #id 

## Functionality

The App plots the county data on the following:
```{r kable, results="asis", echo=FALSE}
var1<-c("Population", "Change", "Population per mile", "Under 5","Under 18",
 "65 and Over", "Percent White", "Percent Black" , "Percent Hispanic", "Percent Asian", "Median Income", "Below Poverty")
desc<-c("Population by County, 2013 estimate", "Population by County, percent change - April 1, 2010 to July 1, 2013", "Population per square mile by county, 2010", "Persons under 5 years, percent by county","Persons under 18 years, percent by county", "Persons 65 years and over, percent by county", "% White by county", "% Black by county", "% Hispanic by county", "% Asian by county",  "Median household income, 2008-2012 by county", "Persons below poverty level, percent, 2008-2012 by county")
kable(cbind(var1,desc))
```

--- .class #id 

## Example Output

Here is the map resulting from selecting "Below Poverty"

```{r plot, echo=FALSE , warning=FALSE}
library(maps)
library(mapproj)
counties <- read.table("http://quickfacts.census.gov/qfd/download/DataSet.txt", header=TRUE, sep=",")
#remove state and US totals
counties<-subset(counties, fips%%1000!=0)
# select poverty
var<-counties$PVY020212
titlein<-"Persons below poverty level, percent, 2008-2012 by county"
  
# generate vector of fill colors for map
    shades <- colorRampPalette(c("yellow", "red"))(5)
    
  
# cut the data into 5 categories and assign colors
  cutpts<-(quantile(var, probs = seq(0,1,.2)))
  cats <- as.integer(cut(var, cutpts, 
  include.lowest = TRUE, ordered = TRUE))
  fills <- shades[cats]

  # plot choropleth map
  map("county", fill = TRUE, col = fills, 
    resolution = 0, lty = 0, projection = "polyconic", 
    myborder = 0, mar = c(0,0,0,0))
  
  # overlay state borders
  map("state", col = "white", fill = FALSE, add = TRUE,
    lty = 1, lwd = 1, projection = "polyconic", 
    myborder = 0, mar = c(0,0,0,0))
  
  # add title
  title(titlein)
  
  # add a legend
  legend.txt<-paste0("between ", signif(cutpts[1:5],3), " and ", signif(cutpts[2:6],3))

  legend("bottomleft", 
    legend = legend.txt, 
    fill = shades[c(1:5)],
    cex = 0.8,
    bty = "n")
```
